Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-08-16 prop_letter.tex
prop_letter.tex Updated from Gerardo...

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality~\cite{BarrettKent2004,Appleby_2005}. The outcome
of a theory that is contextual depends upon
whether  compatible sets of observables are measured together or
separately; a non-contextual theory gives the same results in either
case.  A classic example of the varying opinions on the impact of
imprecision is the disagreement between
Meyer~\cite{PhysRevLett.83.3751} and 
Mermin~\cite{Mermin1999}  on whether finite-precision
measurements invalidate the spirit of the Kochen-Specker
theorem~\cite{kochenspecker1967,peres1995quantum}.  The Kochen-Specker  
theorem is in essence a
mathematical statement about contextuality, asserting that
in a Hilbert space of dimension $d \ge 3$, it is
impossible to associate definite measured values, $\mu(P_i)= 0$ or
$1$, with every projection operator~$P_i$, in such a way that, if a
set of commuting $P_i$ satisfies $\sum_i P_i = 1$, then
$\sum_i \mu(P_i) = 1$.   Meyer showed that one {\it can\/}
assign definite values $0$ or $1$ to the measurement outcome
$\mu(P_i)$ when the Hilbert space is defined over the field of
rational numbers, therefore nullifying the
theorem~\citep{PhysRevLett.83.3751}.  However, Mermin then argued
that measurement outcomes must depend smoothly on slight changes in
the experimental configuration, leading him to assert that the
impact of Meyer's negative result is
``unsupportable''~\citep{Mermin1999}.  The question of
whether contextuality is consistent with experimental
outcomes continues to be an active topic of discussion~\citep{Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,Peres2003,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

@@

A principal result is that recasting the Kochen-Specker theorem in
terms of ``$\delta$-determinism'', with $\delta$ denoting a
\emph{range} of outcomes, reducing to the purely deterministic
Kochen-Specker case for $\delta=0$, can resolve the question of effect
of finite precision and imperfections on contextuality.  A remarkable
consequence of our analysis is that not only does the
quantum-mechanical requirement for contextuality continue to hold,
thus maintaining the Kochen-Specker result for a \emph{non-vanishing}
range of $\delta$, but that at a \emph{certain fixed value} of
$\delta$, there is a sharp transition, parallel in its consequences
to, say, a phase transition, to \emph{non-contextuality}; that is,
quantum mechanics suddenly fails to hold, and all variables can be
specified simultaneously, turning the theory effectively into a
classical theory that can be specified consistently by hidden
variables.  These results provide a new insight into the
Meyer-Mermin debate by presenting a theory in which there is a
\emph{parameter} interpolating between what appeared previously to be
irreconcilable interpretations of contextuality.

>>>>

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality~\cite{BarrettKent2004,Appleby_2005}. The outcome
of a theory that is contextual depends upon
whether  compatible sets of observables are measured together or
separately; a non-contextual theory gives the same results in either
case.  A classic example of the varying opinions on the impact of
imprecision is the disagreement between
Meyer~\cite{PhysRevLett.83.3751} and 
Mermin~\cite{Mermin1999}  on whether finite-precision
measurements invalidate the spirit of the Kochen-Specker
theorem~\cite{kochenspecker1967,peres1995quantum}.  The Kochen-Specker  
theorem is in essence a
mathematical statement about contextuality, asserting that
in a Hilbert space of dimension $d \ge 3$, it is
impossible to associate definite measured values, $\mu(P_i)= 0$ or
$1$, with every projection operator~$P_i$, in such a way that, if a
set of commuting $P_i$ satisfies $\sum_i P_i = 1$, then
$\sum_i \mu(P_i) = 1$.   Meyer showed that one {\it can\/}
assign definite values $0$ or $1$ to the measurement outcome
$\mu(P_i)$ when the Hilbert space is defined over the field of
rational numbers, therefore nullifying the
theorem~\citep{PhysRevLett.83.3751}.  However, Mermin then argued
that measurement outcomes must depend smoothly on slight changes in
the experimental configuration, leading him to assert that the
impact of Meyer's negative result is
``unsupportable''~\citep{Mermin1999}.  The debate on
whether contextuality is consistent with experimental
outcomes continues to be an active topic of research~\citep{Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,Peres2003,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

@@

A principal result is that recasting the Kochen-Specker theorem in
terms of ``$\delta$-determinism'', with $\delta$ 
quantifying the effect of finite-precision uncertainties on measurement 
outcomes, reducing to the purely deterministic
Kochen-Specker case for $\delta=0$, can resolve the question of effect
of finite precision and imperfections on contextuality.  A remarkable
consequence of our analysis is that not only does the
quantum-mechanical requirement for contextuality continue to hold,
thus maintaining the Kochen-Specker result for a \emph{non-vanishing}
range of $\delta$, but that at a \emph{certain fixed value} of
$\delta$, there is a sharp transition, parallel in its consequences
to, say, a phase transition, to \emph{non-contextuality}; that is,
quantum mechanics suddenly fails to hold, and all variables can be
specified simultaneously, turning the theory effectively into a
classical theory that can be specified consistently by hidden
variables.  These results provide a new insight into the
Meyer-Mermin debate by presenting a theory in which there is a
\emph{parameter} interpolating between what appeared previously to be
irreconcilable interpretations of contextuality.
